---
title: "Analysis of Groundedness, RT facilitation and Surprisal Difference"
output: html_document
date: "2024-01-31"
---

```{r}
library(lmerTest)
library(lme4)
library(brms)
library(ggmcmc)
library(mcmcplots)
library(rstanarm)
library(RColorBrewer) # needed for some extra colours in one of the graphs
library(bayesplot)
library(ggplot2)
library(dplyr)
library(tidybayes)
library(modelr)
library(tidyverse)
```


```{r}
dataset_with_error <- read_csv("./final_v2_all_with_error_info.csv",
                   col_types = cols(Condition_ID=col_factor(levels=c("no", "wrong", "correct")),
                                    POS = col_factor(levels = c("Closed","Open")))
                   ) %>% filter(correctness != 'unavailable') 
```



# Surprisal Difference(different types of surprisals) & groundedness

```{r}
dataset_correct <- dataset_with_error %>% filter(correctness!='wrong')
dat_grounding <- 
  dataset_correct %>%
  filter(Condition_ID != "no") %>%
  group_by(WordToken,Subject_ID,Group,Condition_ID) %>%
  summarize(POS=POS[1],gpt2_surp=mean(gpt2_surp),kosmos2_surp=mean(kosmos2_surp),Frequency=mean(Frequency),Length=mean(Length),groundedness=mean(groundedness),RT=mean(RT))
dat_grounding_by_word_token <-
  dat_grounding %>%
  group_by(WordToken,Group,Condition_ID) %>%
  summarize(POS=POS[1],gpt2_surp=mean(gpt2_surp),kosmos2_surp=mean(kosmos2_surp),Frequency=mean(Frequency),Length=mean(Length),groundedness=mean(groundedness),RT=mean(RT)) %>%
  mutate(#groundedness=(10+groundedness)/2,
         kosmos2_gpt2_surp_diff=kosmos2_surp-gpt2_surp,
         Condition=ifelse(Condition_ID=="wrong","Wrong Image Preview","Correct Image Preview"),
         `Part of Speech`=POS,
         contingent_surp=ifelse(Condition_ID=="wrong",gpt2_surp,kosmos2_surp))

model_surp_diff <- lmer(kosmos2_gpt2_surp_diff ~ 1 + `Part of Speech`:groundedness +  Length + Frequency + (`Part of Speech`:groundedness +  Length + Frequency | Group) , filter(dat_grounding_by_word_token, Condition_ID=='correct'))

summary(model_surp_diff)
```


# Surprisal Difference (Same type) & groundedness

```{r}
dat_grounding <- 
  dataset_correct %>%
  filter(Condition_ID != "wrong") %>%
  group_by(WordToken,Subject_ID,Group,Condition_ID) %>%
  summarize(POS=POS[1],kosmos2_surp=mean(kosmos2_surp),Frequency=mean(Frequency),Length=mean(Length),groundedness=mean(groundedness),RT=mean(RT))

dat_grounding_by_word_token <-
  dat_grounding %>%
  group_by(WordToken,Group,Condition_ID) %>%
  summarize(POS=POS[1],kosmos2_surp=mean(kosmos2_surp),Frequency=mean(Frequency),Length=mean(Length),groundedness=mean(groundedness),RT=mean(RT)) # %>%
  #pivot_wider(names_from = Condition_ID, values_from = kosmos2_surp) ##%>%
  #mutate(#groundedness=(10+groundedness)/2,
         #kosmos2_surp_diff= wrong-correct,
         #`Part of Speech`=POS)

dat_grounding_by_word_token_correct <- dat_grounding_by_word_token %>% filter(Condition_ID=='correct') %>% select(-Condition_ID) 
dat_grounding_by_word_token_no <- dat_grounding_by_word_token %>% filter(Condition_ID=='no') %>% select(-Condition_ID)

```

```{r}
combined_df <- inner_join(dat_grounding_by_word_token_correct, dat_grounding_by_word_token_no, by = c("WordToken","Group","POS","groundedness","Frequency","Length")) %>%
  mutate(kosmos2_surp_diff = kosmos2_surp.x - kosmos2_surp.y) %>%
  mutate(RT_diff = RT.x - RT.y) %>%
  select(-starts_with("kosmos2_surp.")) %>%
  select(-starts_with("RT."))
```

```{r}
ggplot(combined_df,aes(x=groundedness,y=kosmos2_surp_diff,color=POS)) +
  geom_point() +
  stat_smooth(method="lm") + 
  theme_bw() +
  xlim(c(-5,10)) +
  ylim(c(-20,7)) +
  xlab("Groundedness of Word in Correct Image") +
  ylab("KOSMOS-2 surp diff(corrrect - no)") +
  theme(legend.position = "top") #+
  #facet_grid(.~Condition) 
```


```{r}
model_surp_diff_same <- lmer(kosmos2_surp_diff ~ 1 + POS:groundedness +  Length + Frequency + (POS:groundedness +  Length + Frequency | Group) , filter(dat_grounding_by_word_token, Condition_ID=='correct'))
```


# RT facilitation & surprisal, groundedness


```{r}
dataset_with_error$RT <- as.numeric(dataset_with_error$RT)
all_correct <- dataset_with_error %>% filter(correctness=="correct")
rt_word_token_by_condition1 <- 
  all_correct %>%
  group_by(WordToken,groundedness,POS,Condition_ID, Group, Frequency, Length) %>%
  summarize(RT=mean(RT)) %>%
  pivot_wider(names_from=Condition_ID,values_from=RT) %>%
  mutate(`RT Diff to Wrong Image`=wrong-correct,
         `RT Diff to No Image`= no-correct) %>%
  pivot_longer(cols=c(`RT Diff to Wrong Image`,`RT Diff to No Image`),names_to="RT Diff to",values_to="RT_diff")

rt_word_token_by_condition2 <- 
  all_correct %>%
  group_by(WordToken,groundedness,POS,Condition_ID, Group, Frequency, Length) %>%
  summarize(kosmos2_surp=mean(kosmos2_surp)) %>%
  pivot_wider(names_from=Condition_ID,values_from=kosmos2_surp, names_prefix = "kosmos2surp_") %>%
  mutate(`Surp Diff to Wrong Image`=kosmos2surp_wrong-kosmos2surp_correct,
         `Surp Diff to No Image`=kosmos2surp_no-kosmos2surp_correct) %>%
  pivot_longer(cols=c(`Surp Diff to Wrong Image`,`Surp Diff to No Image`),names_to="Surp Diff to",values_to="Surp_diff")

combined_df <- inner_join(rt_word_token_by_condition1,rt_word_token_by_condition2)

combined_df$`Surp Diff to` <- factor(combined_df$`Surp Diff to`)

combined_df_wrong <- combined_df %>% filter(`Surp Diff to`=="Surp Diff to Wrong Image")
model_wrong <- glmer(RT_diff ~ Surp_diff + groundedness:POS + Frequency + Length + (Surp_diff + groundedness:POS + Frequency + Length | Group) + (1 | WordToken), data=combined_df_wrong)
summary(model_wrong)

combined_df_no <- combined_df %>% filter(`Surp Diff to`=="Surp Diff to No Image")
model_no <- glmer(RT_diff ~ Surp_diff + groundedness:POS + Frequency + Length + (Surp_diff + groundedness:POS + Frequency + Length | Group) + (1 | WordToken), data=combined_df_no)
summary(model_no)
```
# Error rate analysis 

```{r}
dataset_with_error_avg <- dataset_with_error %>% 
  group_by(WordToken,Condition_ID,Group,correctness) %>%
  summarize(blip2_surp=mean(blip2_surp)) %>%
  group_by(Condition_ID,Group,correctness) %>%
  summarize(blip2_surp=mean(blip2_surp)) %>%
  group_by(Condition_ID,correctness) %>%
  summarize(blip2_surp_avg=mean(blip2_surp),blip2_surp_se = sd(blip2_surp)/sqrt(n())) %>%
  mutate(Condition=spelled_out_conditions[as.character(Condition_ID)]) %>%
  filter(Condition_ID!='no')
  
```
```{r}
dataset_with_error_avg
dataset_with_error_avg <- dataset_with_error_avg %>% rename(`Correctness Status of Words`=correctness)
dataset_with_error_avg
```
```{r}
my_dodge <- position_dodge(0.9)  
ggplot(dataset_with_error_avg,aes(x=Condition,y=blip2_surp_avg,fill=`Correctness Status of Words`)) +
    geom_bar(stat="identity", 
           position=my_dodge,color = "black") +
  geom_errorbar(aes(ymin=blip2_surp_avg-blip2_surp_se, ymax=blip2_surp_avg+blip2_surp_se), width=.2,
                 position=position_dodge(.9)) +
  ylab("BLIP2_surprisal ± Standard Error") +
  ylim(c(0,12)) +
  #scale_fill_manual(values = c("orange", "green"))
  theme_bw() +
  theme(legend.position = c(0.5,0.93),
        legend.direction="horizontal")
#ggsave("img/gpt2_error.pdf",height=4,width=4)
```

# Error prediction model, Correct Vs wrong condition data

```{r}
dataset_here <- dataset_with_error %>% filter(Condition_ID!='no')
dataset_here$Condition_ID <- relevel(dataset_here$Condition_ID, ref = "correct")
dataset_here$POS <- relevel(dataset_here$POS, ref = "Open")
dataset_here$Condition_ID <- droplevels(dataset_here$Condition_ID)
contrasts(dataset_here$Condition_ID) <- "contr.sum"
contrasts(dataset_here$POS) <- "contr.sum"

#m_errors <- brm(correctness ~ Condition_ID*POS + blip2_surp + (Condition_ID | Subject_ID) + (Condition_ID | Group) + (Condition_ID | WordToken), data=dataset_here, family = "bernoulli",warmup = 1000, iter = 4000,cores = 6, chains = 2, control = list(adapt_delta = 0.85),seed = 123)
```

# Bigger brms model causes convergence issues

```{r}
m_errors <- readRDS("saved_models/m_errors.RDS")
```

```{r}
m_errors_glmer <- glmer(correctness ~ Condition_ID*POS + kosmos2_surp + Frequency + Length + (Condition_ID*POS | Group) + (Condition_ID | Subject_ID) + (Condition_ID | WordToken) + (Condition_ID | Word), data=dataset_here, family="binomial",control = glmerControl(optimizer = "bobyqa"))
```



```{r}
summary(m_errors)
```

# Correct Vs no

```{r}

```

# No Vs Wrong 

I would like to know if there's a better way to do the above analysis without running 3 models 
```{r}

```
